Make any stupid plot you want

### 2021-02-01 draft/sections/implementation.tex
veen rereading the aos and hunter papers

To generate figure~\ref{fig:code_scatter_line} in our prototype, we set up a canvas \mintinline{python}{fig} and an axes to draw on \mintinline{python}{ax} and 
\begin{multicols*}{2}
\begin{minted}[linenos]{python}
    fig, ax = plt.subplots()
    artist = Point(table, transforms)
    ax.add_artist(artist)
\end{minted}
\columnbreak
\begin{minted}[linenos]{python}
    fig, ax = plt.subplots()
    artist = Line(table, transforms)
    ax.add_artist(artist)
\end{minted}
\end{multicols*}
 and then pass in \dsection=\mintinline{python}{table} and \vchannel=\mintinline{python}{transforms} into equivalence classes \vartisteq=\mintinline{python}{Point} and  \vartisteq=\mintinline{python}{Line}.

>>>>

To generate the images in figure~\ref{fig:code_scatter_line}, we use the figure \mintinline{python}{fig}  artist and the axes \mintinline{python}{ax} artist that lies inside the figure as the canvas for our graphic.

\begin{multicols*}{2}
\begin{minted}[linenos]{python}
    fig, ax = plt.subplots()
    artist = Point(table, transforms)
    ax.add_artist(artist)
\end{minted}
\columnbreak
\begin{minted}[linenos]{python}
    fig, ax = plt.subplots()
    artist = Line(table, transforms)
    ax.add_artist(artist)
\end{minted}
\end{multicols*}
 We then add to the axes our prototype artist equivalance classes \vartisteq=\mintinline{python}{Point} and  \vartisteq=\mintinline{python}{Line}. We pass in a section of the data \dsection=\mintinline{python}{table} and  paramterize the artist using the \vchannel=\mintinline{python}{transforms} dictionary.

### 2021-02-01 draft/sections/math.tex
veen rereading the aos and hunter papers

\note{should this be in 3rd person passive?}\\
Visualization is generally thought of as structure preserving maps from data into graphics, and in this section we formally define that structure and how it is preserved via equivariant maps. We can then specify that a faithful visual mapping is structure preserving, and apply these constraints to visualizations we may want to develop or implement. We model the data, visual characteristic, and graphic stages of visualization, shown in figure~\ref{fig:artist_stages}, as topological structures that encapsulate types of variables and continuity; by doing so we can develop implementations that keep track of both in ways that let us distribute computation while still allowing assembly and dynamic update of the graphic.

@@

\subsubsection{Variables: Fiber Space \dfiber}
\label{sec:data_fiber}
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the components of the fiber to variable names and types. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_n\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema binding variable name \fname\ to its datatype \ftype. A dataset with multiple variables has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\ftype_1}\times \ldots \ftotal_{\ftype_i} \ldots\times \ftotal_{\ftype_n}
\end{equation}
which is equivalent to 
\begin{equation}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}

@@

\subsubsection{Measurement Scales: Monoid Actions}
\label{sec:data_monoid}
After specifying \dfiber\, we next describe the ways in which we can transform the values by identifying the monoid actions \monoid\ on the \dfiber. We use monoids as the abstraction because they encode composibilty, which maps well to the data transformation process in a software library \cite{yorgeyMonoidsThemeVariations}.

@@

\subsubsection{Continuity: Base Space $K$} 
\label{sec:data_base}
The advantage of fiber bundles is they provide a way to encode the continuity in a dataset as the base space \dbase\ without making assumptions as to what that continuity is. In turn this representation of continuity can then be used to keep track of how the data fits together, for example if a visualization of a very large dataset calls for parallelization.

@@

which is a decomposition of \dfiber. The \dbase\ remains the same because the connectivity of records does not change just because there are fewer elements in each record.

\begin{figure}[ht!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_1k.png}
        %% add box around neighboring P and Map
        \caption{}
        \label{fig:base_example_discrete}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2k.png}
        \caption{}
        \label{fig:base_example_continuous}
    \end{subfigure}
    \caption{These two datasets have the same (time, temperature) fiber. In figure~\ref{fig:base_example_line} the total space \dtotal\ is discrete over points $\dbasepoint \in \dbase$, meaning the records in the fiber are also discrete. In figure~\ref{fig:base_example_plane} \dtotal\ lies over the continuous interval \dbase, meaning the records in the fiber are sampled from a continuous space. 
    \note{revamp figure: F=Plane, k1 = dots, k2=line}}
    \label{fig:base_example}
\end{figure}

@@

\subsubsection{Data: Sections \dsection}
\label{sec:data_section}
While the fiber and base space describe the general structure of all data that lives in the fiber bundle, the sections $\dsection: \dbase\rightarrow \dtotal$ define the datasets that live in the fiber. We generalize Spivak's description of the section as a table of records \cite{spivakSIMPLICIALDATABASES} to any sort of structured dataset such that 
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
such that there is always a map $\pi(\dsection(\dbasepoint)) = \dbasepoint$. There can be many sections \dsection; the space of global sections is $\Gamma(\dtotal)$. For a trivial fiber bundle, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. Because we can decompose the bundle and the fiber, we can formulate $\dsection$ as 
\begin{align}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{align}
where each section $\dsection_i$ is a variable or set of variables.

@@

@@

\subsubsection{Idealized Display \gfiber}
\label{sec:graphic_fiber}
The fiber \gfiber\ is an idealized infinite resolution version of the target display space, for example a 2D screen or 3D printer. In this work, we assume a 2D opaque image $\dfiber=\reals^5$ with elements 
\begin{equation}
(x,\, y,\, r,\, g,\, b) \in \gfiber
\end{equation}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be $\dfiber=\reals^{7}$ such that $(x, y, z, r, g, b, a) \in \gfiber$ specifies the target display. The location coordinates $x$ and $y$ are defined in terms of the display, while the $(r,g,b)$ values are filled in via a lookup on \gbase.

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:graphic_base}
An assumption of graphical representations is that they match the continuity of the data\cite{tufteVisualDisplayQuantitative2001,friendlyBriefHistoryData2008}, but the underlying topology \gbase\ of a graphic may need more dimensions than the data topology \dbase\ so that the glyph can be defined in \dfiber. Therefore we define the base space mapping from graphic \gbase to data \dbase 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
        \end{tikzcd}
\end{equation}

@@

\subsubsection{Renderable Glyph \gsection}
\label{sec:graphic_section}
%pick a pixel which is the bounding box
A section $\gsection: \gbase\rightarrow \gtotal$ defines a piece of the graphical representation of the data. Evaluated on a single \gbasepoint\, \gsection\ returns a single element in \gfiber\. For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{xy}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. Integrating over this region on \gbase\
\begin{align}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align}
yields the color of the pixel $p$.

@@

\subsection{Artist}
\label{sec:artist}
The artist is the function that converts data into graphics; its name is taken from the analogous part of Matplotlib\cite{hunterArchitectureOpenSource} that builds visual elements to pass off to the renderer. The artist \vartist\ is a mapping from \dtotal\ padded with data from $\mathcal{J}(\dtotal)$ to a graphic that is a section \gsection\ in  $\Gamma(\gtotal)$
\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \Gamma(\gtotal) \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
with an intermediate fiber bundle \vtotal\ to hold visual representations and stages
\begin{enumerate}
    \item \vindex\ binding the continuity in the graphic to the continuity in the data (~\ref{sec:graphic_base})
    \item \vchannel\ conversion of data into visual characteristics (~\ref{sec:artist_nu})
    \item \vmark\ assembly of visual variables into a glyph (~\ref{sec:artist_q})
\end{enumerate}

@@

\subsubsection {Visual Fiber Bundle \vtotal}
The visual fiber bundle (\vtotal, \dbase, $\pi$, \vfiber) has section $\vsection: \vtotal \rightarrow \dbase$ that resolves to a visual variable \cite{bertinIIPropertiesGraphic2011,munznerMarksChannels} in fiber \vfiber. The fiber space \vfiber\ is defined in terms of the parameters of the visualization specification- for example aesthetics  in ggplot \cite{wickhamGgplot2ElegantGraphics2016a}, channels in vega\cite{satyanarayanDeclarativeInteractionDesign2014} or parameters in VTK\cite{hanwellVisualizationToolkitVTK2015} and Matplotlib.

@@

\subsubsection{Visual Channels}
\label{sec:artist_nu}
\begin{figure}
    \includegraphics[width=\textwidth]{figures/math/equivariance_nu.png}
    \caption{In this artis, \vchannel\ maps the strings to the emojis. For \vchannel\ to be equivariant, a shuffle in the words should have an equivalent shuffle in the emojies, and a shuffle in the emojies should have an equivalent shuffle in the words.}
    \label{fig:artist_nu}
\end{figure}
As introduced in section~\ref{sec:intro_visual_variables}, there are many ways to encode data visually. We define the visual transformers \vchannel\ as the set of independent conversion functions 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
where $\vchannel_i: \dsection_i \mapsto \vsection_i$ is an equivariant map such that there is a monoid homomorphism from $\dfiber_{i}$ to $vfiber_{i}$. A validly constructed \vchannel\ is one where the  diagram of the monoid transform $m$
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
    \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
commutes such that $\vchannel_i(m_{\delement}(\dtotal_i)) = m_{\velement}(\vchannel_i(\dtotal_i))$. This equivariance constraint yields guidance on what makes for an invalid transform. For example, the conversion $\vchannel_{i}(x) = .5$ does not commute under translation monoid action $t(x) = x+2$  
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + \vchannel(2)\\
    .5 &\neq .5 + .5
\end{align}

@@

>>>>

%% whats the right model/abstraction for a visualization library? 
%% why? the role this math is serving - static type checking/ reflection
%% tracking/automatically inferring
%% bookend w/ how this math does the why 
%% if we deleted this, what would we lose? 
%% can express continuous functions & not just discrete points
\note{should this be in 3rd person passive?}\\
Visualization is generally thought of as structure preserving maps from data into graphics, and in this section we formally define that structure and how it is preserved via equivariant maps. We can then specify that a faithful visual mapping is structure preserving, and apply these constraints to visualizations we may want to develop or implement. We model the data, visual characteristic, and graphic stages of visualization, shown in figure~\ref{fig:artist_stages}, as topological structures that encapsulate types of variables and continuity; by doing so we can develop implementations that keep track of both in ways that let us distribute computation while still allowing assembly and dynamic update of the graphic.

@@

\subsubsection{Variables: Fiber Space \dfiber}
\label{sec:data_fiber}
%%% so we can do units, static + dynamic type checking, so that we can tak about variables 
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the components of the fiber to variable names and types. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_n\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema binding variable name \fname\ to its datatype \ftype. A dataset with multiple variables has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\ftype_1}\times \ldots \ftotal_{\ftype_i} \ldots\times \ftotal_{\ftype_n}
\end{equation}
which is equivalent to 
\begin{equation}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}

@@

\subsubsection{Measurement Scales: Monoid Actions}
\label{sec:data_monoid}
%% pin down the structure we're trying to preserve, this is what we mean by equivarance, most general useful formalism is a monoid, most specifically, more commonly it's a group but sometimes it's a mpnoid
After specifying \dfiber\, we next describe the ways in which we can transform the values by identifying the monoid actions \monoid\ on the \dfiber. We use monoids as the abstraction because they encode composibilty, which maps well to the data transformation process in a software library \cite{yorgeyMonoidsThemeVariations}.

@@

\subsubsection{Continuity: Base Space $K$} 
\label{sec:data_base}
% scatter and line, heatmap express the continuity of the data, deeply implicit in every visualization system but equivariance of continuity of the records, explicit  topplogy/structure preserving - contuinity is the topology   
The advantage of fiber bundles is they provide a way to encode the continuity in a dataset as the base space \dbase\ without making assumptions as to what that continuity is. In turn this representation of continuity can then be used to keep track of how the data fits together, for example if a visualization of a very large dataset calls for parallelization.

@@

which is a decomposition of \dfiber. The \dbase\ remains the same because the connectivity of records does not change just because there are fewer elements in each record.
\begin{figure}[ht!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_1k.png}
        %% add box around neighboring P and Map
        \caption{}
        \label{fig:base_example_discrete}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2k.png}
        \caption{}
        \label{fig:base_example_continuous}
    \end{subfigure}
    \caption{These two datasets have the same (time, temperature) fiber. In figure~\ref{fig:base_example_line} the total space \dtotal\ is discrete over points $\dbasepoint \in \dbase$, meaning the records in the fiber are also discrete. In figure~\ref{fig:base_example_plane} \dtotal\ lies over the continuous interval \dbase, meaning the records in the fiber are sampled from a continuous space. 
    \note{revamp figure: F=Plane, k1 = dots, k2=line}}
    \label{fig:base_example}
\end{figure}

@@

\subsubsection{Data: Sections \dsection}
\label{sec:data_section}
%% how do we represent data
While the fiber and base space describe the general structure of all data that lives in the fiber bundle, the sections $\dsection: \dbase\rightarrow \dtotal$ define the datasets that live in the fiber. We generalize Spivak's description of the section as a table of records \cite{spivakSIMPLICIALDATABASES} to any sort of structured dataset such that 
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
such that there is always a map $\pi(\dsection(\dbasepoint)) = \dbasepoint$. There can be many sections \dsection; the space of global sections is $\Gamma(\dtotal)$. For a trivial fiber bundle, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. Because we can decompose the bundle and the fiber, we can formulate $\dsection$ as 
\begin{align}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{align}
where each section $\dsection_i$ is a variable or set of variables.

@@

%% here are some of our real data structures and here is how it works: dataframe, movies, etc...
%% all the same, special versions of fiber bundles

@@

\subsubsection{Idealized Display \gfiber}
\label{sec:graphic_fiber}
%% bookends sentences
%% first we need a  target output abstraction 
%% this allows us to w/ some level of detail specify a target display
The fiber \gfiber\ is an idealized infinite resolution version of the target display space, for example a 2D screen or 3D printer. In this work, we assume a 2D opaque image $\dfiber=\reals^5$ with elements 
\begin{equation}
(x,\, y,\, r,\, g,\, b) \in \gfiber
\end{equation}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be $\dfiber=\reals^{7}$ such that $(x, y, z, r, g, b, a) \in \gfiber$ specifies the target display.

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:graphic_base}
%% structure preservation on the graphics side 
An assumption of graphical representations is that they match the continuity of the data\cite{tufteVisualDisplayQuantitative2001,friendlyBriefHistoryData2008}, but the underlying topology \gbase\ of a graphic may need more dimensions than the data topology \dbase\ so that the glyph can be defined in \dfiber. Therefore we define the base space mapping from graphic \gbase to data \dbase 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
        \end{tikzcd}
\end{equation}

@@

\subsubsection{Renderering \gsection}
\label{sec:graphic_section}
% couple nu + Q , visual encodings 
% pick a pixel which is the bounding box
A section $\gsection: \gbase\rightarrow \gtotal$ defines a piece of the graphical representation of the data. Evaluated on a single \gbasepoint\, \gsection\ returns a single element in \gfiber\. For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{xy}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. Integrating over this region on \gbase\
\begin{align}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align}
yields the color of the pixel $p$.

@@

%% pdf/svg/etc are different ways of implementing rhos

\subsection{Artist}
\label{sec:artist}
%%% line things up such that they're an equivariant maps 
%%% make a little clearer what E' is 
The artist is the function that converts data into graphics; its name is taken from the analogous part of Matplotlib\cite{hunterArchitectureOpenSource} that builds visual elements to pass off to the renderer. The artist \vartist\ is a mapping from \dtotal\ padded with data from $\mathcal{J}(\dtotal)$ to a graphic that is a section \gsection\ in  $\Gamma(\gtotal)$
%%% \Gamma is total
%%% H is point wise -
%%% this map means we cna do this pointwise 
%%% explain the arrows, chase through here
%%% new disagram w/ section

\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
%%% Q \circ \xi  \nu = \rho, and nu is on tau, and sections are weaker 
with an intermediate fiber bundle \vtotal\ to hold visual representations and stages
\begin{enumerate}
    \item \vindex\ binding the continuity in the graphic to the continuity in the data (~\ref{sec:graphic_base})
    \item \vchannel\ conversion of data into visual characteristics (~\ref{sec:artist_nu})
    \item \vmark\ assembly of visual variables into a glyph (~\ref{sec:artist_q})
\end{enumerate}

@@

\subsubsection {Visual Fiber Bundle \vtotal}
%% structure preserving on the visual side/target space, visualization library 
%% equivariant, latent space of visual parameters, intermediate 
%% don't formally need, let it more cleanly pin down the the fiber equivarance , expresisng the concept that is already in literature,

The visual fiber bundle (\vtotal, \dbase, $\pi$, \vfiber) has section $\vsection: \vtotal \rightarrow \dbase$ that resolves to a visual variable \cite{bertinIIPropertiesGraphic2011,munznerMarksChannels} in fiber \vfiber. The fiber space \vfiber\ is defined in terms of the parameters of the visualization specification- for example aesthetics  in ggplot \cite{wickhamGgplot2ElegantGraphics2016a}, channels in vega\cite{satyanarayanDeclarativeInteractionDesign2014} or parameters in VTK\cite{hanwellVisualizationToolkitVTK2015} and Matplotlib.

@@

\subsubsection{Visual Channels}
\label{sec:artist_nu}
%% equivariant maps 
\begin{figure}
    \includegraphics[width=\textwidth]{figures/math/equivariance_nu.png}
    \caption{In this artis, \vchannel\ maps the strings to the emojis. For \vchannel\ to be equivariant, a shuffle in the words should have an equivalent shuffle in the emojies, and a shuffle in the emojies should have an equivalent shuffle in the words.}
    \label{fig:artist_nu}
\end{figure}
As introduced in section~\ref{sec:intro_visual_variables}, there are many ways to encode data visually. We define the visual transformers \vchannel\ as the set of independent conversion functions 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
where $\vchannel_i: \dsection_i \mapsto \vsection_i$ is an equivariant map such that there is a monoid homomorphism from $\dfiber_{i}$ to $vfiber_{i}$. A validly constructed \vchannel\ is one where the  diagram of the monoid transform $m$
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
    \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
commutes such that $\vchannel_i(m_{\delement}(\dtotal_i)) = m_{\velement}(\vchannel_i(\dtotal_i))$. This equivariance constraint yields guidance on what makes for an invalid transform. For example, the conversion $\vchannel_{i}(x) = .5$ does not commute under translation monoid action $t(x) = x+2$  
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + \vchannel(2)\\
    .5 &\neq .5 + .5
\end{align}

@@

\subsubsection{+ union}
\begin{figure}
    \caption{components of axes + glyphs }
\end{figure}
disjoin union of fibers so we can have an axes/real figure, - diagram operator w/o overplotting, library won't auto stack, have to bake that into data, haskell data/depth, 
hover, mouse over curve, picked a point selecting a little k, + operator/disjoint union in different k, same D
%% but what if it;s supposed to be Bertin invariant? will do w/ inclusion of K
